---
title: "Administrative boundaries"
output:
  html_document:
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{css, echo=FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
```

```{r general options, include = FALSE}
knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  prompt = function(before, options, envir) {
    options(prompt = if (options$engine %in% c("sh", "bash")) "$ " else "> ")
  })

knitr::opts_chunk$set(margin = TRUE, message = FALSE, warning = FALSE,
                      prompt = FALSE, comment = "##", collapse = FALSE,
                      cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
                      fig.width = 4.24725, fig.retina = 2, fig.align = "center")

options(width = 263)

l <- "en_US.UTF-8"
Sys.setenv(LANGAGE = l)
Sys.setlocale(locale = l)
Sys.setlocale("LC_MESSAGES", l)
```

One source of administrative boundaries is [GADM](https://gadm.org). You can
visit the website and download manually the data by country, administrative
level, and format. Two R format are available: `sp` and `sf`. Let's downlaod
all the administrative levels for Vietnam, in `sf` format, and put them into
the `raw data` folder:

```{r eval = FALSE}
download.file("https://biogeo.ucdavis.edu/data/gadm3.6/Rsf/gadm36_VNM_0_sf.rds",
              "../raw data/gadm36_VNM_0_sf.rds")
download.file("https://biogeo.ucdavis.edu/data/gadm3.6/Rsf/gadm36_VNM_1_sf.rds",
              "../raw data/gadm36_VNM_1_sf.rds")
download.file("https://biogeo.ucdavis.edu/data/gadm3.6/Rsf/gadm36_VNM_2_sf.rds",
              "../raw data/gadm36_VNM_2_sf.rds")
download.file("https://biogeo.ucdavis.edu/data/gadm3.6/Rsf/gadm36_VNM_3_sf.rds",
              "../raw data/gadm36_VNM_3_sf.rds")
```

Let's have a look:

```{r}
library(sf)
library(stringr)
library(dplyr) # safer to load last
```

```{r}
gadm0 <- readRDS("../raw data/gadm36_VNM_0_sf.rds")
gadm1 <- readRDS("../raw data/gadm36_VNM_1_sf.rds")
gadm2 <- readRDS("../raw data/gadm36_VNM_2_sf.rds")
gadm3 <- readRDS("../raw data/gadm36_VNM_3_sf.rds")
```

They are basically S3 data frames:

```{r}
gadm3
```

of which the geometry can be extracted and plotted independently of any
associated variable values:

```{r}
plot(st_geometry(gadm1))
```

Let's now clean and tune a bit these data. Cleaning first:

```{r}
gadm1 %<>% 
  select(-NAME_0, -GID_0, -NL_NAME_1, -CC_1) %>% 
  tibble::remove_rownames() %>% 
  mutate_if(is.character, ~ str_trim(.) %>% str_squish())
```

```{r}
gadm2 %<>% 
  select(-NAME_0, -GID_0, -starts_with("NL"), -CC_2) %>% 
  tibble::remove_rownames() %>% 
  mutate_if(is.character, ~ str_trim(.) %>% str_squish())
```

```{r}
gadm3 %<>%
  select(-NAME_0, -GID_0, -starts_with("NL"), -CC_3, -HASC_3) %>% 
  tibble::remove_rownames() %>% 
  mutate_if(is.character, ~ str_trim(.) %>% str_squish())
```

What I want to do now is to add the English names as well as the types, in
Vietnamese and English, of the levels 1 and 2 to `gadm2` and `gadm3` when
missing. Let's first add the English names of provinces as well as the level 1
types, in Vietnamese and English, to `gadm2`:

```{r}
gadm2 <- gadm1 %>% 
  as.data.frame() %>% 
  select(GID_1, VARNAME_1, TYPE_1, ENGTYPE_1) %>% 
  left_join(gadm2, ., "GID_1") %>% 
  select(GID_1, NAME_1, VARNAME_1, TYPE_1, ENGTYPE_1, GID_2, NAME_2, VARNAME_2, TYPE_2, ENGTYPE_2, HASC_2)
```

```{r}
gadm3 <- gadm2 %>% 
  as.data.frame() %>% 
  select(-GID_1, -NAME_1, -NAME_2, -geometry) %>%
  left_join(gadm3, ., "GID_2") %>% 
  select(GID_1, NAME_1, VARNAME_1, TYPE_1, ENGTYPE_1, GID_2, NAME_2, VARNAME_2,
         TYPE_2, ENGTYPE_2, HASC_2, GID_3, NAME_3, VARNAME_3, TYPE_3, ENGTYPE_3)
```

And let's now put these cleaned and tuned data to the `cleaned data` folder:

```{r eval = FALSE}
saveRDS(gadm0, "../clean data/gadm0.rds")
saveRDS(gadm1, "../clean data/gadm1.rds")
saveRDS(gadm2, "../clean data/gadm2.rds")
saveRDS(gadm3, "../clean data/gadm3.rds")
```



